---
title: "R Basics"
author: "Mia Kupres"
format:
  html:
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

## Addition
```{r}
2 + 2
```
## Multiplication and Addition
```{r}
5 * 5 + 2
```
# division and subtraction
```{r}
5 / 5 - 3
```
# log, exponential, pi
```{r}
log(exp(pi))
```

# sinusoids
```{r}
sin(pi/2) 
```
# power
```{r}
exp(1)^(-2) 
```
# square root
```{r}
sqrt(8)
```
# sequences
```{r}
1:5
```
# sequences
```{r}
seq(1, 10, by=2)
```
# repeat 2 three times

```{r}
rep(2, 3) 
```
## Next, we’ll use assignment to make some objects:
# put 1 + 2 in object x
```{r}
x <- 1 + 2 
```
# same as above with fewer keystrokes
```{r}
x = 1 + 2
```
# same
```{r}
1 + 2 -> x
```
# view object x
```{r}
x
```

# put 9 times 3 in y and view the result
```{r}
(y = 9 * 3)
```

# put 5 standard normals into z and print z
```{r}
(z = rnorm(5))
```

## It is worth pointing out R’s recycling rule for doing arithmetic. In the code below,c() [concatenation] is used to create a vector. Note the use of the semicolon for multiple commands on one line.
```{r}
x = c(1, 2, 3, 4); y = c(2, 4, 6, 8); z = c(10, 20); w = c(8, 3, 2)
```

# 1*2, 2*4, 3*6, 4*8
```{r}
x * y
```

# 1+10, 2+20, 3+10, 4+20
```{r}
x + z
```
# what happened here?
```{r}
y + w 
```
# To work your objects, use the following commands:
```{r}
ls() # list all objects
```
# list every object that contains "my"
```{r}
ls(pattern = "m") 
```
# list of available data sets
```{r}
data()
```
# remove object "dummy"
```{r}
rm(dummy) 
```
# remove almost everything (use with caution)
```{r}
rm(list=ls()) 
```
# html help and documentation
```{r}
help.start() 
```
# specific help (?exp is the same)
```{r}
help(exp) 
```
# get working directory
```{r}
getwd()
```
# change working directory
```{r}
setwd()
```
# end the session (keep reading)
```{r}
q()
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```