##### In this section, we give a brief introduction on using R for time series. We assume that astsa has been loaded. To create a time series object, use the command ts. Related commands are as.ts to coerce an object to a time series and is.ts to test whether an object is a time series. First, make a small data set:
```{r}
(mydata = c(1, 2, 3, 2, 1)) #make it and view it
```
##### Now make it a time series object:
```{r}
(myts = as.ts(mydata)) #make it a time series and view it
```
##### Make it a time series that starts in 1950 :
```{r}
(myts = ts(mydata, start=1950)) #make it a time series starting in 1950 and view it
```
##### Now make it a quarterly time series that starts in 1950-III:
```{r}
#make it a quarterly time series starting in 1950-III and view it
(myts = ts(mydata, start=c(1950,3), frequency=4)) 
time(myts) #view the sampled times
```
##### Next, we’ll look at lagging and differencing. First make a simple series, xt:
```{r}
xt = ts(1:5)
xt
```
##### Now, column bind (cbind) lagged values of xt and you will notice that lag(x) is forward lag, whereas lag(x, -1) is backward lag.
```{r}
cbind(xt, lag(xt), lag(xt, -1))
```
##### Compare bind and ts.intersect:
```{r}
ts.intersect(xt, lag(xt), lag(xt, -1))
```
##### To difference a series, ∇t = xt − xt−1, use
```{r}
diff(xt)
```
##### But note that diff(xt, 2) is not second order differencing, it is xt − xt−2. For second order differencing, that is,  $\nabla^2 xt$, do one of these: diff(diff(xt))
```{r}
diff(xt, 2)#xt − xt−2
diff(diff(xt))#second order differencing, that is,  $\nabla^2 xt$
diff(xt, diff=2) # same thing, and so on for higher order differencing.
```
##### We will also make use of regression via lm(). First, suppose we want to fit a simple linear regression, $y = \alpha + \beta x + \epsilon$. In R, the formula is written as y~x:
```{r}
set.seed(1999)
x = rnorm(10)
y = x + rnorm(10)
summary(fit <- lm(y~x) )
```
```{r}
plot(x, y)
abline(fit)
```

##### All sorts of information can be extracted from the lm object, which we called fit. For example,
```{r}
resid(fit) # will display the residuals
fitted(fit) # will display the fitted values
lm(y ~ 0 + x) # will exclude the intercept
```
##### You have to be careful if you use lm() for lagged values of a time series. If you use lm(), then what you have to do is align the series using ts.intersect. Please read the warning Using time series in the lm() help file [help(lm)].

##### Here is an example regressing astsa data, weekly cardiovascular mortality (cmort) on particulate pollution (part) at the present value and lagged four weeks (part4). First, we create ded, which consists of the intersection of the three series:

```{r}
library(astsa)
data(cmort)
exists("cmort")
```

```{r}
ded = ts.intersect(cmort, part, part4=lag(part,-4))
```
##### Now the series are all aligned and the regression will work.
```{r}
summary(fit <- lm(cmort~part+part4, data=ded, na.action=NULL) )
```

##### There was no need to rename lag(part,-4) to part4, it’s just an example of what you can do. There is a package called dynlm that makes it easy to fit lagged regressions. The basic advantage of dynlm is that it avoids having to make a data frame; that is, line 2 would be avoided.

##### In Problem 2.1, you are asked to fit a regression model
$$
x_t = \beta t + \alpha_1 Q_1(t) + \alpha_2 Q_2(t) + \alpha_3 Q_3(t) + \alpha_4 Q_4(t) + w_t
$$
##### where xt is logged Johnson & Johnson quarterly earnings (n = 84), and $Q_i(t)$ is the indicator of quarter i = 1, 2, 3, 4. The indicators can be made using factor.

```{r}
trend = time(jj) - 1970 # helps to `center' time
Q = factor(cycle(jj) ) # make (Q)uarter factors
reg = lm(log(jj)~0 + trend + Q, na.action=NULL) # no intercept
model.matrix(reg) # view the model design matrix
```

```{r}
summary(reg) # view the results (not shown)
```

##### The workhorse for ARIMA simulations is arima.sim. Here are some examples; no output is shown here so you’re on your own.
```{r}
x = arima.sim(list(order=c(1,0,0), ar=.9), n=100) + 50 # AR(1) w/mean 50
```
```{r}
x = arima.sim(list(order=c(2,0,0), ar=c(1,-.9)), n=100) # AR(2)
```
```{r}
x = arima.sim(list(order=c(1,1,1), ar=.9 ,ma=-.5), n=200) # ARIMA(1,1,1)
```
##### An easy way to fit ARIMA models is to use sarima from astsa. The script is used in Chapter 3 and is introduced in Section 3.8.